---
title: "Water Jugs Problem"
author: "Wisnu Adi Nurcahyo"
date: "September 15th, 2018"
output:
  html_document:
    df_print: paged
---

An AI example of water jugs problem in R using BFS (Breadth-First Search) strategy.
This R Notebook project is converted to HTML which means we won't be able to run the R code below on our browser.
You need an RStudio to runs them interactively.

## Problem Description

There are two jugs with a maximum capacity of $N$ and $M$ consecutively.
Let say that the first water jug $A$ has a capacity of $N$ liters.
The second water jug $B$ has a capacity of $M$ liters.
Suppose that we want to fill either one of them to about $X$ liters where $0 \leq X \leq \max(N, M)$. Otherwise, there is no solution. Please not that $N$ and $M$ must follows $N, M \geq 0$.

What is the **fastest way** to get the desired $X$? Assume that we have an unlimited amount of supply.

As an example, we have jug $A$ with a capacity of $12$ ($N = 12$).
Also, a jug $B$ with a capacity of $11$ ($N = 11$).
We want to get a jug of $6$ liters ($X = 6$).
What is the correct step to get there?

```{r}
# Jug A capacity
N <- 12

# Jug B capacity
M <- 11

# Desired result
X <- 6
```

## Production Rules

We could say that "production rule" is the same as the "next possible move".
Hence, we need to determine the production rules for calculating its _branching factors_.
This is important as it is our way to deciding searching startegy.
Since we need to find the **fastest way**, we need a **complete** and **optimal** startegy.
BFS is one of them.

Below is our production rules. Our initial state is $(0, 0)$ in the form of pair $(P, Q)$.
Since we have $X = 6$, our goal state is either $(6, Q)$ or $(P, 6)$.
We also could write it as $(X, Q)$ or $(P, X)$.

* Fill the jug $A$ by a tap. That's means $(12, Q)$ since $A$ has a capacity of $12$.
* Fill the jug $B$ by a tap. That's means $(P, 11)$ since $B$ has a capacit of $11$.
* Throw away water in the jug $A$. That's means $(0, Q)$.
* Throw away water in the jug $B$. That's means $(P, 0)$.
* Fill the jug $A$ by jug $B$.
  That's means $(P + Q, 0)$ if $P + Q \leq N$.
  Otherwise, $(P + Y, Q - Y)$ where $Y = N - P$.
* Fill the jug $B$ by jug $A$.
  That's means $(0, Q + P)$ if $Q + P \leq M$.
  Otherwise, $(P - Y, Q + Y)$ where $Y = M - Q$.

From those possible moves, our branching factors is $6$.

```{r}
# Starting point
S <- c(0, 0)

# User defined production rules as possible moves
possible_moves <- c(
  (function(p, q) c(12, q)),
  (function(p, q) c(p, 11)),
  (function(p, q) c(0, q)),
  (function(p, q) c(p, 0)),
  (function(p, q) if ((p + q) <= N) c(p + q, 0) else c(p + (N - p), q - (N - p))),
  (function(p, q) if ((q + p) <= M) c(0, q + p) else c(p - (M - q), q + (M - q)))
)
```

## Breadth-First Search Implementation

The implementation is not really different from the usual BFS.
However, we need to store the traced solution nodes.
That's means we will use backtracking method here.
Well, not really.
I just will use tree data structure instead.

I assumed that the reader understand about BFS, backtracking, and tree data structure so I won't explain it again here.
First, we need to implement a function that return a string as a key for our list.

```{r}
to_node <- function(x) paste0("(A = ", x[[1]], ", B = ", x[[2]], ")")
```

Then, we implements our BFS algorithm here.

```{r}
library(data.tree)

solution <- Node$new(to_node(S))

bfs <- function() {
  queue <- list()
  queue <- c(queue, list(S))

  visited <- list()

  childs <- list()
  childs[[to_node(S)]] <- solution

  while (length(queue) > 0) {
    current <- queue[[1]]
    queue <- queue[-1]

    node <- to_node(current)
    visited[[node]] <- TRUE

    p <- current[[1]]
    q <- current[[2]]

    if (p == X || q == X) {
      queue <- list()
      next
    }

    for (next_move in possible_moves) {
      next_move <- next_move(p, q)
      next_node <- to_node(next_move)

      if (is.null(visited[[next_node]])) {
        queue <- c(queue, list(next_move))
        visited[[node]] <- TRUE

        childs[[next_node]] <- childs[[node]]$AddChild(next_node)
      }
    }
  }
}
```

Now, run the BFS and print the solution.
We will print its entire state tree.

```{r, fig.height = 15}
# Run our BFS first
bfs()

# Styling our Graph plot
SetGraphStyle(solution, rankdir = "TB")
SetEdgeStyle(solution, arrowhead = "vee", color = "grey35", penwidth = 2)

SetNodeStyle(solution, style = "filled, rounded", shape = "box", fontcolor = "Black", fillcolor = "LightBlue", tooltip = GetDefaultTooltip)
Do(solution$leaves, function(node) SetNodeStyle(node, fontcolor = "White", fillcolor = "YellowGreen"))

# Display it
plot(solution)
```

We have retrieved the solution.
Take a look on the YellowGreen node.
Find either $A$ or $B$ that contains $6$.
It is the solution.

## Closing

I'm new in R.
This is my first R code that pushed on GitHub.
I will learn how to visualize the data in the nicer way.
My code might not efficient.
However, I successfully implemented the BFS algorithm with tree data structure for storing its solution.
Before closing this page, verify our solution on the [Water Jugs Puzzle][1] game.
Choose the highest level which is level 6.
Then, follow the solution that we have retrieved.
Enjoy!

[1]: https://www.mathsisfun.com/games/jugs-puzzle.html
