---
title: "Simulated Annealing"
author: "Wisnu Adi Nurcahyo"
date: "September 20th, 2018"

output:
  html_notebook:
    css: presentation.css
---

## Problem Description

Find the **minimum value** of the following $f(x_1, x_2)$ function where $-10 \leq x_1, x_2 \leq 10$
by using **simulated annealing** algorithm.

$$
f(x_1, x_2) = -\left\lvert{\sin(x_1) \cos(x_2) \exp\biggl(\left\lvert{1 - \frac{\sqrt{{x_1}^2 + {x_2}^2}}{\pi}}\right\rvert\biggr)}\right\rvert
$$

## Implementation

We will implements the simulated annealing algorithm in R.
The visualization (by `plotly`) also included.

```{r message = FALSE}
library(plotly)
```


### The Function $f(x_1, x_2)$ in R

```{r}
f <- (function(x1, x2)
  -abs(sin(x1) * cos(x2) * exp(abs(1 - (sqrt((x1 ** 2) + (x2 ** 2)) / pi))))
)
```


### Initializing Contour Plot

In this report, I'm learning about how to visualize my data.
As the getting started, we will see the initial contour plot below.
How? First, we need a function to generate a (matrix) mesh grid.
It is a two dimensional matrix. Then, fill it up by using $f(x_1, x_2)$ above.

```{r}
# A function for generating mesh grid
meshgrid <- function(x1, x2 = x1) {
  n <- length(x1)
  m <- length(x2)
  
  X1 <- matrix(rep(x1, each = m), nrow = m, ncol = n)
  X2 <- matrix(rep(x2, each = n), nrow = m, ncol = n)
  
  return(list(X1, X2))
}
 
# The generated mesh grid
seq <- seq(-10.0, 10.0, by = 0.1)
mesh_grid <- meshgrid(seq)
mesh_grid_x1 <- mesh_grid[[1]]
mesh_grid_x2 <- mesh_grid[[2]]
 
# Re-initialize `mesh_grid` variable
mesh_grid <- mesh_grid_x1
dim <- dim(mesh_grid)
 
# Fill up `mesh_grid` by using f(x1, x2)
for (i in 1:dim[[1]])
  for (j in 1:dim[[2]])
    mesh_grid[[i, j]] <- f(mesh_grid_x1[[i, j]], mesh_grid_x2[[i, j]])
 
# Display the contour plot
plot_ly(
  x = seq,
  y = seq,
  z = mesh_grid,
  type = "contour",
  hoverinfo = "none",
  contours = list(coloring = "heatmap")
) %>% colorbar(title = "Value of f(x1, x2)") %>% layout(
  xaxis = list(title = "x1", showticklabels = TRUE),
  yaxis = list(title = "x2", showticklabels = TRUE)
)
```

The contour plot above is the generated value from all $x_1$ and $x_2$ where $-10 \leq x_1, x_2 \leq 10$ in $f(x_1, x_2)$ function. The solution is nowhere in the plot.
Since we want to find its **minimum value**, we could predict that the solution is on the dark purple colored area.

### Simulated Annealing Implementation

```{r}
# Initialize random solution
start <- list(x1 = -2, x2 = 5, temperature = 1000)
start <- c(start, list(acceptance = f(start$x1, start$x2)))
 
# Initialize current solution
current <- start
best <- c(current, acceptance = f(current$x1, current$x2))
 
# Generate random point (x1, x2)
generate <- (function() runif(1, -10, 10))

# Initialize a new mesh grid for z = f(x, y)
mesh_grid_x <- c()
mesh_grid_y <- c()
mesh_grid_z <- c()
 
# Main simulated annealing algorithm
while (current$temperature > 0) {
  # Generate a new random solution
  newly <- list(x1 = generate(), x2 = generate())
  newly <- c(newly, acceptance = f(newly$x1, newly$x2))
  
  # Energy changes a.k.a \Delta E
  changes <- newly$acceptance - current$acceptance
  
  if (changes < 0) {
    current$x1 <- newly$x1
    current$x2 <- newly$x2
    
    current$acceptance <- newly$acceptance
    
    # Adding a new value for each mesh grid
    mesh_grid_x <- c(mesh_grid_x, current$x1)
    mesh_grid_y <- c(mesh_grid_y, current$x2)
    mesh_grid_z <- c(mesh_grid_z, current$acceptance)
    
    if (newly$acceptance < best$acceptance) {
      best$x1 <- newly$x1
      best$x2 <- newly$x2
      
      best$acceptance <- newly$acceptance
    }
  }
  else {
    # Considering move to the current (x1, x2)
    P <- exp((-changes) / current$temperature)
    R <- runif(1)
    
    if (R < P) {
      current$x1 <- newly$x1
      current$x2 <- newly$x2
    }
  }
  
  # Decrease the temperature randomly
  current$temperature <- current$temperature - runif(1, 0.0001, 0.01)
}

# Retrieve the solution and then print it
solution <- list(x1 = best$x1, x2 = best$x2, value = best$acceptance)
print(solution)
```

By using simulated annealing above, we got the solution. It is $(-8.049689, 9.661877)$ in the form of $(x_1, x_2)$.
The value of $f(x_1, x_2)$ is $-19.20814$.
Remember that the solution could be a bit different if we start over and run the algorithm again.
However, the result would be around $-19$.

### Display the Contour Plot

```{r}
plot_ly(
  x = mesh_grid_x,
  y = mesh_grid_y,
  z = mesh_grid_z,
  type = "contour",
  hoverinfo = "none",
  contours = list(coloring = "heatmap")
) %>% colorbar(title = "Value of f(x1, x2)") %>% layout(
  xaxis = list(title = "x1", showticklabels = TRUE),
  yaxis = list(title = "x2", showticklabels = TRUE)
)
```

We all know that the solution is around $-19$ and it is produced by (around) $(8.0, 9.7)$.
Hence, by looking at the contour plot above, the solution is on the that dark purple color.
